---
title: "Modeling and prediction for movies"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
library(gridExtra)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `movies`. Delete this note when before you submit 
your work. 

```{r load-data}
load("movies.Rdata")
```



* * *

## Part 1: Data

* Generalizability : Following the description of 'Codebook for Movies dataset', this dataset is randomly sampled, which means that it has generalizability.
* Causality : As this dataset is observed information, there is no random assignment, which means it doesn't have causality.

* * *

## Part 2: Research question

* **Can we guess audience score from some features of movies such as genre, runtime?**

  - I believe the success of movies is dependent on audience's reaction.I heard 'Rotten Tomatoes' is well known for the honest reviews from audience.
  - Can we find out any relationship between audience score and features like
    1. genre : Genre of movie
    2. runtime : Runtime of movie(in minutes)
    3. thtr_rel_year : Year the movie is released in theaters
  - and if the movie is good, do people tend to visit websites to vote?
    4. imdb_num_votes : Number of votes on IMDB

* * *

## Part 3: Exploratory data analysis

* Let's find out the range or groups of the features.

**1. audience score**

```{r}
movies %>% select(audience_score) %>% summarise(min = min(audience_score)
                                                ,mean = mean(audience_score)
                                                ,max = max(audience_score))
hist(movies$audience_score, main = "Histogram of Audience Score")
```

* The range of audience score is between 11 and 100. From the website, the score is given in percent. It makes sense the scores is under 100.


**2. runtime**
```{r}
movies %>% filter(!is.na(runtime)) %>% select(runtime) %>% summarise(min = min(runtime)
                                                                     , mean = mean(runtime)
                                                                     , max = max(runtime))
```

**3. thtr_rel_year**
```{r}
movies %>% select(thtr_rel_year) %>% summarise(min = min(thtr_rel_year), max = max(thtr_rel_year))
```



### Visualization

```{r}
par(mfrow = c(1,3))
hist(movies$audience_score, col = 'skyblue', main = 'Histogram of Audience Score')
hist(movies$thtr_rel_year, col = 'pink', main = 'Histogram of released year in theater')
hist(movies$runtime, col = 'orange', main = 'Histogram of runtime')
```

- The distribution of audience score and released year in theater are a little bit left skewed.
- Most movies have the runtime between 50 and 150 minutes.

**4. imdb_num_votes**
```{r}
ggplot(movies, aes(imdb_num_votes)) + geom_area(stat='bin') +  
ggtitle("Count of Number of votes on IMDB") +  
theme(plot.title = element_text(hjust = 0.5))
```

- The number of voters are mainly less than 125,000.

**5. genre**
```{r}
movies %>% select(genre) %>% group_by(genre) %>% summarise()

```


```{r}
genre_bar <- movies %>% select(genre)
ggplot(genre_bar, aes(x = genre)) + geom_bar(fill='grey') +  
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5 )) +  
  ggtitle("Bar chart of Genre")
```

- There are 11 kinds of genre (not exactly 11, there are 'Others'). The largest area is drama.

```{r}
a = ggplot(movies, aes(audience_score, runtime)) + geom_point(color = 'skyblue')
b = ggplot(movies, aes(audience_score, thtr_rel_year)) + geom_point(color = 'orange')
c = ggplot(movies, aes(audience_score, imdb_num_votes)) + geom_point(color = 'pink')
grid.arrange(a,b,c, nrow = 3, ncol=1)
ggplot(movies, aes(thtr_rel_year, runtime)) + geom_point(color = 'green')
```

- It doesn't look like there are collinearity between these variables.
- But the movies with high score might have large number of voters.

```{r}
ggplot(movies, aes(audience_score, genre)) + geom_violin(scale='area', color = 'grey')
```

- We can see documentary and art genre tend to have good scores from audience.

* * *

## Part 4: Modeling


```{r}
# Exclude null data in 'runtime' variable.
movies_2 <- movies %>% select(audience_score, genre, runtime, thtr_rel_year, imdb_num_votes) %>%  
  filter(!is.na(runtime))

# Linear model
movie_mlr = lm(audience_score ~ genre + runtime + thtr_rel_year + imdb_num_votes, data = movies_2)

summary(movie_mlr)

# 1. backward
backward <- step(movie_mlr, direction = 'backward', trace = FALSE)
backward$anova

# 2. forward
forward <- step(movie_mlr, direction = 'forward')
forward$anova

# 3. Both
both <- step(movie_mlr, direction = 'both', trace = 0)
both$anova
```

* Through backward elimination method, only 'runtime' variable shows higher AIC score. The forward selection chose no variable so final model is below.

```{r}
final <- lm(audience_score ~ runtime, data = movies_2)
```

* Let's diagnose the model.
```{r}
plot(final$residuals)
plot(final$residuals - movies_2$runtime)
hist(final$residuals)
qqnorm(final$residuals)

```

* The model does not have perfectly normal residuals with mean 0 and the variability seems constant.

```{r}
summary(final)
```

* For each minute increase in runtime, the movie score from audience is expected to be higher on average by 0.19.
* The R-squared value is very low, 0.03 so this model is not good model to see the correlationship between audience score and other features.


* * *

## Part 5: Prediction

* Reference for the prediction : https://www.rottentomatoes.com/m/moonlight_2016
* Following the reference, the movie 'Moonlight' has 111 minutes runtime.

```{r}
new_runtime <- data.frame(runtime=c(111))
predict(final, newdata = new_runtime)

```

* The actual score from audience is 79%.
* Let's calculate the prediction interval.

```{r}
summary(final)
t= qt(0.025, df = nrow(movies_2)-2)
high = 0.19 + t*0.04
low = 0.19 - t*0.04
```

* We are 95% confident that, the model predicts that runtime slope is between 0.11 and 0.27.

* * *

## Part 6: Conclusion

* The result of prediction is not close to the real value.
* Following the model and diagnosis, the model uses only one feature 'runtime' among what I selected. It seems suspicious that runtime can predict the score. In this context, R-squared and Adjusted R-squared are low. Overall, the features that I selected are not appropriate to predict the score from audience.
* Probably the variable 'best_pic_nom', 'best_pic_win' or top200_box can be related to the score but I excluded because it is too obvious.
* Audience tend to evaluate movies based on its content. You can see this from the violin plot between genre and score.
* If we have information related to content, it would be a better model.
